A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-22 latex/main.tex
Minor tweaks

% P3. Clarifies relationships between pango X lineages, simplifying
% the Omicron sublineage.
% 
% SHZ: I don't think knowing the relationships between Pango Xs is what's really important.
% Sure, phylogeneticists would get a kick out of it,
% but I don't think the majority of Cell readership would.
% I think what is important is knowing
% how many emerging recombinants carrying new combinations of mutations of concern
% we are actually dealing with in real time.
% Clarifying the relationships between Pango Xs make this possible.

@@

Alignments read from the VCZ Zarr file were first preprocessed
to mask any non nucleotide characters (A,C,G or T) as missing data.
% https://github.com/jeromekelleher/sc2ts-paper/blob/main/inference/viridian_config.toml
Then, we filtered any samples that had missing data at more than 500 sites,
excluding the 100 sites identified as problematic
(this was set to 10,000 sites for the initial period up to 2020-03-01
where sampling density is low).

@@

The LS model \cite{LiStephens2003} is an approximation of
the coalescent with recombination
that captures many of the key features of
the joint processes of mutation and recombination.
It is widely used in a range of applications in population 
genetics~\cite{mcvean2019linkage}.
The LS model is an HMM that models a focal genome
as a sequence of nucleotides (the observed states),
which are probabilistically emitted
as an imperfect mosaic of a set of genomes
in a reference panel (the hidden states).
The process of switching between the genomes in the reference panel
is governed by a transition matrix,
and mismatches between a focal genome and a genome of the reference panel
(from which it is ``copied'') are permitted as emissions.

@@

\subsubsection*{Choice of k paramter}
% This is a key decision that we will be referencing the discussion, so 
% deserves its own section

>>>>

% P3. Clarifies relationships between pango X lineages, simplifying
% the Omicron sublineage.
% 
% SHZ: I don't think knowing the relationships between Pango Xs is what's really important.
% Sure, phylogeneticists would get a kick out of it,
% but I don't think the majority of Cell readership would.
% I think what is important is knowing
% how many emerging recombinants carrying new combinations of mutations of concern
% we are actually dealing with.
% Clarifying the relationships between Pango Xs make this possible.

@@

Alignments read from the VCF Zarr file were first preprocessed
to mask any non nucleotide characters (A,C,G or T) as missing data.
% https://github.com/jeromekelleher/sc2ts-paper/blob/main/inference/viridian_config.toml
Then, we filtered any samples that had missing data at more than 500 sites,
excluding the 100 sites identified as problematic
(this was set to 10,000 sites for the initial period up to 2020-03-01
where sampling density is low).

@@

The LS model \cite{LiStephens2003} is an approximation of
the coalescent with recombination
that captures many of the key features of
the joint processes of mutation and recombination.
The LS model is an HMM that models a focal genome
as a sequence of nucleotides (the observed states),
which are probabilistically emitted
as an imperfect mosaic of a set of genomes
in a reference panel (the hidden states).
The process of switching between the genomes in the reference panel
is governed by a transition matrix,
and mismatches between a focal genome and a genome of the reference panel
(from which it is ``copied'') are permitted as emissions.

@@

\subsubsection*{Choice of k parameter}
% This is a key decision that we will be referencing the discussion, so 
% deserves its own section
